---
title: "SFI Data Analysis"
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(shiny)
library(data.table)
library(DT)
library(sfi)
```

This Shiny app provides a graphical user interface for the `sfi` package. Upload an mzML file and set the parameters to perform feature extraction and alignment for Single File Injection (SFI) data. The workflow follows the steps outlined in the package vignette.

```{r, context="ui"}
sidebarLayout(
  sidebarPanel(
    h4("1. Upload Data"),
    fileInput("mzml_file", "Upload mzML File", accept = c(".mzML")),
    
    h4("2. Set Parameters"),
    textInput("qcseq", "QC/Blank", "1,1,0,1,1,0,1,1,0"),
    p(em("Comma-separated list of 1s (QC) and 0s (Blank)."), style = "font-size: 0.9em; color: #666;"),

    numericInput("n_samples", "Number of Samples (n)", 150, min = 1),

    hr(),

    h5("Data Resolution Mode"),
    radioButtons("resolution_mode", "Select Data Type:",
                 choices = list("High Resolution" = "high",
                               "Low Resolution (Unit Mass)" = "low"),
                 selected = "high"),

    conditionalPanel(
      condition = "input.resolution_mode == 'low'",
      numericInput("low_res_deltart", "Delta RT for Low Res", 5, min = 1),
      numericInput("low_res_snr", "Signal-to-Noise Ratio", 3.0, min = 0.1, step = 0.1),
      p(em("Low resolution mode aggregates data by integer mass before peak finding."),
        style = "font-size: 0.9em; color: #666;")
    ),

    hr(),
    
    h5("Alignment Parameters"),
    checkboxInput("optimize_params", "Auto-optimize Parameters?", value = TRUE),

    conditionalPanel(
      "input.resolution_mode == 'low'",
      p(em("Note: Parameter optimization is not available for low resolution data."),
        style = "font-size: 0.9em; color: #888;")
    ),

    conditionalPanel(
      "input.optimize_params == true && input.resolution_mode == 'high'",
      numericInput("getqc_intensity_threshold", "Optimization Intensity Threshold", 1000000, min = 0)
    ),

    numericInput("idelta", "Interval (idelta)", 92, min = 0, step = 0.01),
    numericInput("windows", "Window", 632, min = 1),
    uiOutput("optim_note_ui"),
    
    hr(),
    
    h5("Advanced Parameters"),
    numericInput("deltart", "Delta RT for grouping", 10, min = 1),
    numericInput("minn", "Minimum points for feature", 6, min = 1),
    numericInput("ppm", "PPM Tolerance (ppm)", 3, min = 0),

    hr(),
    
    actionButton("run_analysis", "Run Analysis", icon = icon("play")),
    br(),
    br(),
    h5("Analysis Status"),
    verbatimTextOutput("analysis_message")
  ),
  
  mainPanel(
    tabsetPanel(
      tabPanel("Feature Table", 
               br(),
               p("After analysis, the feature table will be displayed below."),
               downloadButton("download_table", "Download Table (.csv)"),
               hr(),
               DTOutput("feature_table")
      ),

    )
  )
)
```

```{r, context="server"}
# Reactive values to store results
feature_table_data <- reactiveVal(NULL)
analysis_message <- reactiveVal("Ready for analysis.")

observeEvent(input$run_analysis, {
  req(input$mzml_file)
  
  # Reset previous results
  feature_table_data(NULL)
  analysis_message("Starting analysis...\n")

  # Show a notification that analysis is running
  showNotification("Analysis in progress...", duration = NULL, id = "analysis_notification")

  tryCatch({
    # --- 1. Parse qcseq input ---
    qc_seq_vector <- as.numeric(trimws(unlist(strsplit(input$qcseq, ","))))
    if (any(is.na(qc_seq_vector)) || !all(qc_seq_vector %in% c(0, 1))) {
      stop("Invalid QC/Sample Sequence. Please use a comma-separated list of 1s (QC) and 0s (Sample).")
    }

    # --- 2. Load mzML data ---
    analysis_message("Step 1/5: Loading mzML data...\n")
    mzml_data_val <- getmzml(input$mzml_file$datapath)


    # --- 3. Perform peak picking ---
    if (input$resolution_mode == "low") {
      analysis_message("Step 2/5: Performing low resolution 2D peak picking...\n")
      peaklist <- find_peaks_low_res(
        mz = mzml_data_val[,1],
        rt = mzml_data_val[,3],
        intensity = mzml_data_val[,2],
        deltart = input$low_res_deltart,
        snr = input$low_res_snr
      )
    } else {
      analysis_message("Step 2/5: Performing high resolution 2D peak picking...\n")
      peaklist <- find_2d_peaks(mz = mzml_data_val[,1], rt = mzml_data_val[,3], intensity = mzml_data_val[,2])
    }

    idelta_val <- input$idelta
    windows_val <- input$windows

    # --- 4. Parameter Optimization (Optional) ---
    if (input$optimize_params && input$resolution_mode == "high") {
      analysis_message(paste0("Step 3/5: Optimizing parameters with getqc(), using initial guesses window=", windows_val, " and idelta=", idelta_val, "...\n"))
      sfmsub <- peaklist[peaklist$intensity > input$getqc_intensity_threshold, ]

      if (nrow(sfmsub) == 0) {
        stop(paste("No peaks found above the intensity threshold of", input$getqc_intensity_threshold, "for parameter optimization. Try a lower threshold."))
      }

      optimized_params <- getqc(
        mz = sfmsub$mz,
        rt = sfmsub$rt,
        intensity = sfmsub$intensity,
        idelta = idelta_val,
        window = windows_val,
        qcseq = qc_seq_vector,
        n = input$n_samples,
        deltart = input$deltart
      )
      windows_val <- optimized_params[1]
      idelta_val <- optimized_params[2]
      analysis_message(paste0("Step 3/5: Parameters optimized. Using new window = ", round(windows_val, 2), " and idelta = ", round(idelta_val, 2), "\n"))
    } else if (input$resolution_mode == "low" || !input$optimize_params) {
       analysis_message("Step 3/5: Using manually set parameters.\n")
    }

    # --- 5. Align features ---
    analysis_message("Step 4/5: Aligning features with getsfm()...\n")
    sfi_aligned <- getsfm(
      mz = peaklist$mz,
      rt = peaklist$rt,
      intensity = peaklist$intensity,
      idelta = idelta_val,
      windows = windows_val,
      qcseq = qc_seq_vector,
      n = input$n_samples,
      deltart = input$deltart,
      minn = input$minn,
      ppm = input$ppm
    )

    # --- 6. Generate feature table ---
    analysis_message("Step 5/5: Generating final feature table...\n")
    sfi_aligned$peakid <- paste0(round(sfi_aligned$qcmz, 4), '@', round(sfi_aligned$qcrt))
    feature_table <- dcast(setDT(sfi_aligned[, c('sampleidx', 'intensity', 'peakid')]),
                           peakid ~ sampleidx,
                           value.var = "intensity",
                           fun.aggregate = sum)

    # Store the result
    feature_table_data(feature_table)
    
    analysis_message("Analysis successful!\n")
    showNotification("Analysis successful!", type = "message")

  }, error = function(e) {
    showNotification(paste("An error occurred:", e$message), type = "error", duration = 10)
    analysis_message(paste("Error:", e$message, "\n"))
  }, finally = {
    # Remove the notification
    removeNotification("analysis_notification")
  })
})

# Render the data table
output$feature_table <- renderDT({
  req(feature_table_data())
  datatable(feature_table_data(), options = list(scrollX = TRUE, pageLength = 10, search = list(search = "")))
})

# Display messages
output$analysis_message <- renderText({
  analysis_message()
})

# Display optimization note
output$optim_note_ui <- renderUI({
  if(input$optimize_params && input$resolution_mode == "high") {
    p(em("The values for interval and window above will be used as initial guesses for the optimization."), style = "font-size: 0.9em; color: #666;")
  }
})

# Handle download
output$download_table <- downloadHandler(
  filename = function() {
    paste0("feature_table_", Sys.Date(), ".csv")
  },
  content = function(file) {
    fwrite(feature_table_data(), file)
  }
)
